### 2014-12-13 paper.tex
Add roadmap in intro

In this paper, we compare some of the metrics of gerrymandering that have been proposed. They include metrics based on the shape of districts, metrics based on population in the district and the surrounding area, and a recently proposed metric that measures the resulting bias.

@@

  \section{Model}
  In this paper, we will conduct an analysis on the correlation between different measures of compactness and political bias, as a method for determining the overall asymmetry of a given political area.  Furthermore, we will examine the political processes involved in drawing those areas and compare the processes with the levels of gerrymandering exhibited.

@@

  \section{Methods}

@@

  \subsection {Computing Voter Efficiency}
  We calculate voter efficiency using data from a variety of sources.  The US Census Data, along with the Harvard Election Data Archive~\cite{heda} presents a great resource for Federal Election results.  State-by-state elections are best sourced from the respective state agencies that govern the election, such as the Secretary of State's office or Board of Elections.  To calculate voter efficiency, we first go back to the definition of wasted votes.  This measure of voter efficiency will only analyze two-party, single-winner elections.  Thus, for a given party $p$, their opposing party $o$ and a given district $i$, the number of wasted votes  if party $p$ wins ($V_{p} > V_{o}$)  is given by
    \[ V_{wpi} = V_{pi} - \frac{V_{pi}+V_{oi}}{2}\]
  as all votes beyond 50\% are wasted. Otherwise, if party $p$ loses the district, then
    \[V_{wpi} = V_{pi}\]
  as all votes that do not go towards electing a candidate are wasted.

@@

  \section{Analysis}

@@

  \section{Discussion}
  \subsection{Usefulness of Different Metrics}
  \subsubsection{Compactness Metrics}
  As measures for gerrymandering, none of the measures of compactness alone is a perfect measure.  In each case, there were some states that scored poorly on a compactness metric, but for which there was no evidence of partisan gerrymandering.  It is possible that in these cases there was a bipartisan gerrymander, or a gerrymander which did not happen to cause many wasted votes in the election we considered; our data are insufficient to decide.  In addition, our data cannot show a causal relationship.  In particular, while gerrymandering may cause noncompact districts, it could also be the case that states which tend to draw heavily gerrymandered districts also tend to draw noncompact districts.  However, the presence of noncompact districts does seem to generally indicate that there may be gerrymandering, and conversely states with compact districts tended to have low partisan bias.

@@

  \section{Conclusion}
  Partisan gerrymandering damages the fundamental nature of our representative democracy.  It produces unfair advantages that prevent citizens from being represented fairly.  Real redistricting is too complex for a single simple metric to be able to quantify all types of gerrymandering, but by combining multiple metrics, it is possible to catch a number of different types of gerrymanders in a specific and quantifiable way.  Using quantifiable metrics of partisan gerrymandering, we hope to create a standard that can delineate results that are blatantly biased, regardless of any ability to discover the intent of those who created the plans.

>>>>

In this paper, we compare some of the metrics of gerrymandering that have been proposed. They include metrics based on the shape of districts, metrics based on population in the district and the surrounding area, and a recently proposed metric that measures the resulting bias.  In Section~\ref{s:background} we discuss past research on gerrymandering and compactness.  In Section~\ref{s:model} we propose a model for gerrymandering, and in Section~\ref{s:methods} we introduce the data we use to test it.  Finally, in Section~\ref{s:analysis} we present our analysis of that data, and in Section~\ref{s:discussion} we discuss its implications.

@@

  \section{Model}\label{s:model}
  In this paper, we will conduct an analysis on the correlation between different measures of compactness and political bias, as a method for determining the overall asymmetry of a given political area.  Furthermore, we will examine the political processes involved in drawing those areas and compare the processes with the levels of gerrymandering exhibited.

@@

  \section{Methods}\label{s:methods}

@@

  \subsection{Computing Voter Efficiency}
  We calculate voter efficiency using data from a variety of sources.  The US Census Data, along with the Harvard Election Data Archive~\cite{heda} presents a great resource for Federal Election results.  State-by-state elections are best sourced from the respective state agencies that govern the election, such as the Secretary of State's office or Board of Elections.  To calculate voter efficiency, we first go back to the definition of wasted votes.  This measure of voter efficiency will only analyze two-party, single-winner elections.  Thus, for a given party $p$, their opposing party $o$ and a given district $i$, the number of wasted votes  if party $p$ wins ($V_{p} > V_{o}$)  is given by
    \[ V_{wpi} = V_{pi} - \frac{V_{pi}+V_{oi}}{2}\]
  as all votes beyond 50\% are wasted. Otherwise, if party $p$ loses the district, then
    \[V_{wpi} = V_{pi}\]
  as all votes that do not go towards electing a candidate are wasted.

@@

  \section{Analysis}\label{s:analysis}

@@

  \section{Discussion}\label{s:discussion}
  \subsection{Usefulness of Different Metrics}
  \subsubsection{Compactness Metrics}
  As measures for gerrymandering, none of the measures of compactness alone is a perfect measure.  In each case, there were some states that scored poorly on a compactness metric, but for which there was no evidence of partisan gerrymandering.  It is possible that in these cases there was a bipartisan gerrymander, or a gerrymander which did not happen to cause many wasted votes in the election we considered; our data are insufficient to decide.  In addition, our data cannot show a causal relationship.  In particular, while gerrymandering may cause noncompact districts, it could also be the case that states which tend to draw heavily gerrymandered districts also tend to draw noncompact districts.  However, the presence of noncompact districts does seem to generally indicate that there may be gerrymandering, and conversely states with compact districts tended to have low partisan bias.

@@

  \section{Conclusion}\label{s:conclusion}
  Partisan gerrymandering damages the fundamental nature of our representative democracy.  It produces unfair advantages that prevent citizens from being represented fairly.  Real redistricting is too complex for a single simple metric to be able to quantify all types of gerrymandering, but by combining multiple metrics, it is possible to catch a number of different types of gerrymanders in a specific and quantifiable way.  Using quantifiable metrics of partisan gerrymandering, we hope to create a standard that can delineate results that are blatantly biased, regardless of any ability to discover the intent of those who created the plans.
